---
title: Telegram Chats History Analysis
params:
  date: !r Sys.Date()
output:
  pdf_document:
    latex_engine: xelatex
    dev: cairo_pdf
  html_document: default
mainfont: Helvetica
sansfont: Helvetica
mathfont: Consolas
monofont: Consolas
fontsize: 12pt
---
```{r, echo=F, message=F, warning=F}

library(tidyverse)
library(lubridate)
library(dplyr)
library(odbc)
library(stringi)
library(kableExtra)
library(config)

config <- config::get()
con <- DBI::dbConnect(odbc::odbc(),
                        driver = "/usr/local/lib/psqlodbcw.so",
                        database = config$dbname,
                        user = config$dbuser,
                        password=ifelse(is.null(config$dbpassword), '', config$dbpassword),
                        host = config$dbhost,
                        port = config$dbport)
```
```{r useful_functions, echo=F, message=FALSE, warning=FALSE}
#name processing
process_name <- function(name) { 
  name
  #stri_trans_general(name, "russian-latin/bgn") # we used to need this before we used good fonts
}
```

# Chats

We have exported messages from several Telegram chats. How many messages are there in our DB from each chat?  (Limited to a certain pattern of the chat_title)
```{r, echo=F, warning=F}
q <- dbSendQuery(con, "
        SELECT
            c.title,
            m.id
        FROM
            messages m
        INNER JOIN
            chats c
            ON c.id = m.chat_id
        WHERE
            c.title LIKE ?")
dbBind(q, config$chat_title_pattern)
dbtbl <- dbFetch(q)
tbl <- dbtbl %>%
  mutate(chat_title = title) %>%
  group_by(chat_title) %>%
  summarize(message_count = n()) %>%
  select(chat_title, message_count) %>%
  mutate(chat_title = reorder(chat_title, message_count))

```
```{r echo=F, warning=F, fig.asp=0.2}

tbl %>%
  ggplot(aes(chat_title, message_count, fill=chat_title)) + 
  geom_bar(stat="identity") +
  guides(fill="none") +
  scale_x_discrete(name="chat title") +
  scale_y_continuous(name="message count") +
  coord_flip() +
  theme_light(base_family = "Helvetica")
  
sorted <- tbl %>% arrange(desc(message_count))
first <- as.character(sorted$chat_title[1])
second <- as.character(sorted$chat_title[2])
second_msg_count <- sorted$message_count[2]
```

The obvious leader is `r first` chat. `r second` is in the second place having `r second_msg_count` messages. 

# `r first`
This is a chat of opinions, holywar and flame.  Let's find out who are the most opinionated users.  Who posted and how many messages in this chat?  Here and below we'll look mainly at the users who have posted 100 messages or more.  We'll look at message distributions over several parameters: users, day of week, hour of day, message content types and finally by text message length.

```{r echo=F, warning=F, message=F}
fetch_messages_from <- function(chat) {
  q_messages <- con %>% DBI::dbSendQuery("
        SELECT
              u.first_name,
              u.last_name,
              m.send_date,
              m.content_type,
              m.message_text
          FROM
              users u
          INNER JOIN
              messages m
              ON u.id = m.sender_user_id
          INNER JOIN
              chats c
              ON c.id = m.chat_id
          WHERE
              c.title = ?
  ")
  DBI::dbBind(q_messages, chat)
  dbmessages <- DBI::dbFetch(q_messages)
  messages <- dbmessages %>% 
    mutate(name = paste(first_name, last_name))  %>%
    mutate(name = process_name(name)) %>%
    mutate(day_of_week = wday(send_date, week_start = 1, label=T, abbr=F)) %>%
    mutate(hour_of_day = hour(send_date)) %>%
    select(-first_name, -last_name) 
  users <- messages %>% 
    group_by(name) %>%
    summarize(message_count = n()) %>%
    arrange(desc(message_count)) %>%
    mutate(name = reorder(name, message_count)) %>%
    mutate(name = as.character(name))
  
  list(messages, users)  
}
```
```{r echo=F, warning=F, message=F}
result <- fetch_messages_from(first)
messages <- result[[1]]
users <- result[[2]]
```
```{r echo=F, message=F, warning=F, fig.asp = 0.6}
filtered_users <- users %>%
  filter(message_count > 100)

filtered_users %>%
  ggplot(aes(name, message_count)) +
  coord_flip() +
  geom_bar(aes(fill=name), stat="identity") +
  guides(fill="none") + 
  theme(axis.text.y = element_text(size=8), axis.text.x = element_text(size=8)) +
  scale_x_discrete(limits=rev(filtered_users$name)) +
  scale_y_continuous(name="message count") + 
  theme_light(base_family = "Helvetica")
```

## Day of Week

What is the most popular day of week?
```{r echo=F, warning=F, message=F}
tbl <- messages %>%
  mutate(day_of_week = wday(send_date, label=T, abbr=F, week_start=1)) %>%
  group_by(day_of_week) %>% 
  summarize(message_count = n()) %>%
  select(day_of_week, message_count)

```
<div class = "row">
<div class = "col-md-6">
```{r echo=F, warning=F}
tbl %>% 
  kable(col.names=c("day of week", "message count"))  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
</div>
<div class = "col-md-6">
```{r echo=F, message=F, warning=F}
tbl %>%
  ggplot() +
  geom_bar(aes(day_of_week, message_count, fill = day_of_week), stat="identity") +     coord_flip() + 
  guides(fill = "none") +
  scale_x_discrete(limits=rev(tbl$day_of_week)) + 
  xlab("day of week") + 
  ylab("message count")
```
</div>
</div>
`r tbl$day_of_week[1]` and `r tbl$day_of_week[2]` seem to be the heaviest days.  

```{r results="asis", echo=F}
cat('\\pagebreak')
```
Let's see the distribution of messages by people and by day of week. 

```{r messages_users_day_of_week, echo=F, message=F, warning=F}
tbl <- 
  messages %>%
  group_by(name, day_of_week) %>%
  mutate(message_count = n()) %>%
  inner_join(filtered_users, by="name") %>%
  mutate(user_message_count = message_count.y) %>%
  mutate(message_count = message_count.x) %>% 
  select(day_of_week, name, message_count, user_message_count)
  

#options(repr.plot.width = 14, repr.plot.height = 8)

tbl %>%
  ggplot() +
  geom_point(aes(day_of_week, name, size=message_count, color=message_count)) +
  scale_x_discrete(limits=levels(tbl$day_of_week)) +
  scale_y_discrete(limits=rev(filtered_users$name)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(colour=guide_colourbar(title=""),
         size=guide_legend(title="message count")) +
  xlab("day of week") +
  scale_colour_distiller(palette = "Spectral")
```
```{r results='asis', echo=F}
cat('\\pagebreak')
```

## Hour of Day / Day of Week (HOD / DOW)

```{r echo=F, message=F, warning=F}
tbl <- messages %>%
  select(name, day_of_week, hour_of_day) %>% 
  rename(dow=day_of_week, hod = hour_of_day) %>%
  group_by(dow, hod)

tbl %>%
  summarize(count = n()) %>%
  ggplot() +
  geom_point(aes(hod, dow, size=count, colour=count)) +
  scale_x_discrete(limits = seq(0,23)) +
  scale_y_discrete(limits = levels(tbl$dow)) +
  scale_colour_distiller(palette = "Spectral") + 
  guides(size = "none", colour = guide_colourbar(title="message count")) +
  xlab("hour of day") + 
  ylab("day of week")

```

From this chart a simple conclusion is that on Friday the "busy" hours are from 10 to 15, with a peak at 11.   
On Monday we see some activity at 12 and on Tuesday at 14.  
Wednesday has a more even spread of messages from 10-15, with a peak at 15. 

```{r results="asis", echo=F}
cat('\\pagebreak')
```
## Hour of Day
Distribution of messages by hour and user on a given day.  

### Friday
```{r echo=F, message=F, warning=F}
userhod <- function (data, day) {
  data %>%
  filter(dow==day) %>%
  group_by(hod, name) %>%
  summarize(count = n()) %>%
  inner_join(filtered_users, by='name') %>%
  ggplot() +
  geom_point(aes(hod, name, size=count, colour=count)) + 
  scale_colour_distiller(palette="Spectral") + 
  scale_x_discrete(limits = seq(0,23)) +
  scale_y_discrete(limits = rev(filtered_users$name)) + 
  xlab("hour of day")
}

userhod(tbl, "Friday")
```

### Wednesday
```{r echo=F, message=F, warning=F}
userhod(tbl, "Wednesday")
```

## Message Content Types
```{r echo=F, message=F, warning=F}
ctsummary <- messages %>% 
  group_by(content_type) %>%
  summarize(message_count = n()) %>%
  arrange(-message_count) 

ctsummary %>%
  kable() %>%
  kable_styling()
```

The dominating content type is of course `r ctsummary$content_type[1]`   

Let's see what is the distribution of message count by message types and users.  We arrange user list by descending text message count. 

```{r echo = F, message=F, warning=F}
ctsusers <- messages %>%
  group_by(name, content_type) %>%
  summarize(message_count = n()) %>%
  inner_join(filtered_users, by="name") %>%
  rename(user_message_count = message_count.y) %>%
  rename(message_count = message_count.x) %>%
  select(name, content_type, message_count, user_message_count) %>%
  arrange(-user_message_count, -message_count)

```
```{r echo=F, message=F, fig.asp = 1.2} 
ctsusers %>%
  ggplot() +
  geom_point(aes(content_type, name, colour=message_count, size=message_count)) +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) + 
  guides(fill = "none", size = "none") +#guide_legend(title="message count")) + 
  xlab("content type") + 
  scale_colour_distiller(palette = "Spectral") +
  scale_y_discrete(limits = rev(filtered_users$name))
```

We can also explore `messagePhoto` content type in a bit more depth, f.e. let's see who posted how many photo messages arranging from the user with most photo messages:

```{r echo=F, message=F, warning=F}
content_type_plot <- function(ctype) {
  ctypeusers <- ctsusers %>%
    filter(content_type == ctype) %>%
    arrange(desc(message_count))
  
  ctypeusers %>%
    ggplot() +
    geom_bar(aes(name, message_count, fill = name), stat="identity") +
    scale_x_discrete(limits = rev(ctypeusers$name)) +
    coord_flip() +
    guides(fill = "none")
}

content_type_plot('messagePhoto')
```

`messageSticker` content type has yet another distribution:

```{r echo=F, message=F, fig.asp = 0.5}
content_type_plot('messageSticker') 
```

Overall as you see sticker is not the most popular content type.

## Text Message Lengths

Let's see the distribution of message length by person. Again filtering only those who had at least 50 messages.  We use a boxplot here to show the 25th and 75th percentile, and the spread min max as the edges of the line.  The dot represents the mean, the vertical line - the median.  We use color to represent the message count sent by a certain user as another dimension of comparison.

```{r echo=F, message=F, warning=F}

usr_msg_txtlen <- messages %>%
  inner_join(filtered_users) %>%
  filter(content_type=='messageText') %>%
  mutate(text_length = str_length(message_text)) %>%
  group_by(name) %>%
  mutate(mean = mean(text_length)) %>%
  ungroup() %>%
  mutate(name = reorder(name, mean)) %>%
  select(name, text_length, mean, message_count)

usr_msg_txtlen %>%
  ggplot() + 
  geom_boxplot(aes(name, text_length, fill = message_count), outlier.shape=NA) + 
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) +
  scale_y_discrete(limits=seq(0,300,50)) + 
  ylab("message text length") +
  geom_point(aes(name, mean)) +
  coord_flip() +
  scale_fill_distiller(palette = "Spectral")
```

The conclusion to draw here - is that the majority of messages length is well under 100 character, with median below 50 characters. Some users are more verbose, but verbosity does not correlate much with the frequency or total number of messages sent.  People who send a lot of messages don't seem to be very verbose.

```{r results="asis", echo=F}
cat('\\pagebreak')
```

## Conversations

Let's study the structure of conversations.  We consider a single day in a chat to be a single conversation.  We consider all messages before 6am to belong to the previous day - thus let's subtract  6 hours from `send_date` before taking a `date` component from it.  

Here is the distribution of conversations by the number of messages within a conversation:

```{r echo=F, message=F, warning=F}
convos <- messages %>%
  mutate(send_date = send_date - hours(6)) %>%
  mutate(date = date(send_date)) %>%
  group_by(date) %>%
  summarize(n_messages = n()) %>%
  arrange(desc(n_messages)) 

convos_by_days <- convos %>%
  group_by(n_messages) %>%
  summarize(n_convos = n()) %>%
  arrange(desc(n_convos))

total_convos = nrow(convos)
median_convo_length = quantile(convos$n_messages, 0.5)
mean_convo_length = mean(convos$n_messages)
```
```{r echo=F, message=F, warning=F}
convos_by_days %>% 
  filter(n_messages < 105) %>%
  ggplot() + 
  geom_bar(aes(n_messages, n_convos, fill=n_messages), stat="identity") + 
  geom_vline(xintercept=median_convo_length, color="red") +
  geom_vline(xintercept=mean_convo_length, color="green")
```

Median is red and mean is green.  In total there are `r total_convos` convos in this chat. 

### Conversation starters and supporters
There are users who send the first message on a given day.  As the first message we understand a message sent after 6am, all others can be considered to be sent on a previous day.  Thus we want to see the distribution of the first message occurrences by users.  We also want to see who is supporting the discussion - the distribution of the second message occurrences by users. 

```{r echo=F, message=F, warning=F}
starters <- messages %>% 
  mutate(date =  date(send_date)) %>%
  mutate(hour = hour(send_date)) %>%
  filter(hour > 6) %>%
  group_by(date) %>% 
  arrange(desc(date), send_date) 

pairs <- starters %>%
  summarize(first = first(name), second = nth(name, 2))  %>%
  arrange(desc(date))

couples <- pairs %>%
  filter(!is.na(second)) %>%
  group_by(first, second) %>%
  count() %>%
  arrange(-n)

couples %>% kable()
```

We see that a lot of first and second message authors are the same, this is because sometimes a user splits her messages in several, let's collapse and view a sequence of messages from a single author as a single message.  Here is the distribution of conversations by starters and supporters (first and second message author):

```{r echo=F, message=F, warning=F}
filtered_starters <- starters %>%
  filter(row_number() == 1 | name != first(name))

collapsed <- filtered_starters %>%
  summarize(first = first(name), second = nth(name, 2))  %>%
  arrange(desc(date))

collapsed_couples <- collapsed %>% group_by(first, second) %>%
  filter (!is.na(second)) %>%
  count() %>%
  arrange(-n)

collapsed_couples %>% kable()
```


Also the distribution of conversations by first author:

```{r echo=F, message=F, warning=F}

first_author <- pairs %>% 
  group_by(first) %>%
  count() %>%
  arrange(-n)

first_author %>% kable()

```

Distribution of conversation by second author:
```{r echo=F, message=F, warning=F}

second_author <- collapsed %>% 
  group_by(second) %>%
  count() %>%
  arrange(-n)

second_author %>% kable()
```

We also see that some conversations do not have a second author, let's see how many conversations have just a single author and their distribution, and also a % of conversations of a given author where he didn't get any response:

```{r echo=F, message=F, warning=F}
singlers <- collapsed %>%
  filter(is.na(second)) %>%
  group_by(first) %>% 
  count() %>%
  left_join(first_author, by="first")  %>%
  rename(n_single = n.x, n_total=n.y) %>%
  mutate(percent = round(n_single/n_total*100,2)) %>%
  arrange(-n_total) 

singlers %>% kable()
```

